---
title: "Exploratory Assessment of the Downscaled GRACE product"
author: "Tejasvi Hora"
date: "`r Sys.Date()`"
analyst: |
  | Tejasvi Hora
output: 
  html_document:
    df_print: kable
    toc: true
    toc_float: true
    toc_depth: 4
    theme: spacelab
knit: (function(inputFile, encoding) { 
          rmarkdown::render(inputFile,
                            encoding = encoding, 
                            output_file = 'dGWS_analysis_221015.html',  
                            output_dir = '/Users/tejasvi/Dropbox/WB/GRACE_Ensemble/Scripts/Downscaled/') })
editor_options: 
  chunk_output_type: console
---

## Introduction
This report relies on the downscaled GRACE data-provided by __Prof. Di Long, Li Chen and Ximing Wang.__

The purpose of this report is to:

* Exploratory Assessment of the downscaled GRACE-GWS (groundwater storage anomaly)
* Comparison of the downscaled product with in-situ well measurements in India 
* Comparison of the downscaled 0.5 degree product and non-Downscaled interpolated GRACE dataset (the one I had produced)

**Please note:** It appears that the downscaled product now covers entire regions (i.e. SA, MENA and SSA). It will be worth discussing the 
methodology used to move from individual basins to regions.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidync)
library(data.table)
library(tidyverse)
library(lubridate)
library(zyp)
library(sf)
library(raster)
library(terra)
library(RColorBrewer)
library(rasterVis)
library(xts)
library(stars)
library(fasterize)
```

## Load and pre-process the datasets

Main steps involved:

* Loading the dataset: `GWS_mean_05deg.nc`
* Linking points with World Bank regions data

```{r echo = FALSE, collapse = TRUE, cache=TRUE, results='hide'}
proj_dir = "/Users/tejasvi/Dropbox/WB/GRACE_Ensemble/"

pathData = '/Users/tejasvi/Dropbox/gwflagship_GRACEdownscaling/Downscaled TWS_GWS v2/'
gws_mean_05 = 
  tidync(paste0(pathData, 'GWS_mean_05deg.nc')) %>%
  hyper_tibble() %>% as.data.table()

#Create a date sequence
date.seq = 
  seq(as.Date("2003/2/1"), as.Date("2021/9/1"), "month") %>%
  as.data.table() %>%
  rownames_to_column() %>%
  .[, rowname := as.integer(rowname)]
colnames(date.seq) = c('rowname', 'date')

#Merge the dates using the rowname column
gws_mean_05 = 
  gws_mean_05 %>%
  merge(date.seq, by.x = 'time', by.y = 'rowname', all.x = T) %>%
  .[, ':='(yearmon = substr(date, 1, 7),
           GWSA_mean = (GWSA_CLSM+GWSA_Noah)/2)] 
  
#Convert to wide
gws_05_wide = 
  gws_mean_05 %>% 
  dplyr::select(GWSA_mean, lat, lon, date) %>%
  tidyr::spread(key = date, value = GWSA_mean) %>%
  mutate(cell_id = paste0(lon, lat))

#Convert to spatial object
gws_05.v = 
  st_as_sf(gws_05_wide, coords = c("lon", "lat"), 
           crs = "+proj=longlat +datum=WGS84 +no_defs")

#Load the studyregions
med.v = st_read(paste0(pathData, 'studyregions/med.shp'))

sa.v = st_read(paste0(pathData, 'studyregions/sa.shp'))

saf.v = st_read(paste0(pathData, 'studyregions/saf.shp'))

####Load World Regions
wb_regions = 
  st_read(paste0(proj_dir, "Spatial Files/WB_Regions/WB_countries_Admin0_10m.shp")) %>%
  dplyr::select(WB_NAME, ISO_A2, ISO_A3, ISO_N3, TYPE, REGION_WB) %>%
  filter(TYPE != 'Dependency') %>%
  st_make_valid()

#Merge country data with GRACE
gws_05.v = 
  gws_05.v %>%
  st_make_valid() %>%
  st_join(wb_regions)

#Add the WB regions to the data.table version
gws_05_wide = 
  gws_05_wide %>%
  merge(gws_05.v[,c(213:219)] %>% st_drop_geometry(), by = 'cell_id', all.x = T)

```

## Trend Analysis {.tabset .tabset-pills}

The steps involved in assessing trends were:

* Conversion of monthly GWS estimates to annual
* Estimating _linear_ trends by estimating Sen's Slope
* Significance of trend not assessed

### MENA 

```{r echo=FALSE, collapse = TRUE}
#Conduct a trend analysis on each pixel

#First, convert values from monthly to year
gws_annual = 
  gws_05_wide %>% 
  .[as.character(REGION_WB) == 'Middle East & North Africa'] %>%
  data.table::melt(id.vars = c("lon", "lat", "cell_id"),
                   measure.vars = 3:214,
                   variable.name = "ymd", value.name = "GWSA_mean") %>% 
  .[, year := substr(ymd, 1,4)] %>%
  .[, an_mean := mean(GWSA_mean), .(lat, lon, year)] %>% 
  .[,.(lat, lon, cell_id, year, an_mean)] %>% distinct() %>%
  dcast(lon + lat + cell_id ~ year, value.var = "an_mean") %>% .[order(cell_id)]

gws_trends = 
  apply(gws_annual[, 4:ncol(gws_annual)], 1, zyp.yuepilon)

gws_trends_estimates = 
  gws_annual[,1:3] %>%
  mutate(trends = gws_trends[2,], sig = gws_trends[6,]) %>%
  mutate(lon = ifelse(lon>180, lon-360, lon))

gws_trends_estimates = 
  gws_trends_estimates %>%
  mutate(trends = ifelse(trends < -2, -2, trends)) %>%
  mutate(trends = ifelse(trends > 2, 2, trends)) 

gws.raster = rasterFromXYZ(gws_trends_estimates[,.(lon, lat, trends)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     breaks = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5, 2),
     col = color_pal, 
     legend.args = list(text = 'Trend (cm/yr)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```

### SSA

```{r echo=FALSE, collapse = TRUE}

#First, convert values from monthly to year
gws_annual = 
  gws_05_wide %>% 
  .[as.character(REGION_WB) == 'Sub-Saharan Africa'] %>%
  data.table::melt(id.vars = c("lon", "lat", "cell_id"),
                   measure.vars = 3:214,
                   variable.name = "ymd", value.name = "GWSA_mean") %>% 
  .[, year := substr(ymd, 1,4)] %>%
  .[, an_mean := mean(GWSA_mean), .(lat, lon, year)] %>% 
  .[,.(lat, lon, cell_id, year, an_mean)] %>% distinct() %>%
  dcast(lon + lat + cell_id ~ year, value.var = "an_mean") %>% .[order(cell_id)]

gws_trends = 
  apply(gws_annual[, 4:ncol(gws_annual)], 1, zyp.yuepilon)

gws_trends_estimates = 
  gws_annual[,1:3] %>%
  mutate(trends = gws_trends[2,], sig = gws_trends[6,]) %>%
  mutate(lon = ifelse(lon>180, lon-360, lon))

gws_trends_estimates = 
  gws_trends_estimates %>%
  mutate(trends = ifelse(trends < -2, -2, trends)) %>%
  mutate(trends = ifelse(trends > 2, 2, trends)) 

gws.raster = rasterFromXYZ(gws_trends_estimates[,.(lon, lat, trends)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     breaks = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5, 2),
     col = color_pal, 
     legend.args = list(text = 'Trend (cm/yr)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```



### SA

```{r echo=FALSE}

#First, convert values from monthly to year
gws_annual = 
  gws_05_wide %>% 
  .[as.character(REGION_WB) == 'South Asia'] %>%
  data.table::melt(id.vars = c("lon", "lat", "cell_id"),
                   measure.vars = 3:214,
                   variable.name = "ymd", value.name = "GWSA_mean") %>% 
  .[, year := substr(ymd, 1,4)] %>%
  .[, an_mean := mean(GWSA_mean), .(lat, lon, year)] %>% 
  .[,.(lat, lon, cell_id, year, an_mean)] %>% distinct() %>%
  dcast(lon + lat + cell_id ~ year, value.var = "an_mean") %>% .[order(cell_id)]

gws_trends = 
  apply(gws_annual[, 4:ncol(gws_annual)], 1, zyp.yuepilon)

gws_trends_estimates = 
  gws_annual[,1:3] %>%
  mutate(trends = gws_trends[2,], sig = gws_trends[6,]) %>%
  mutate(lon = ifelse(lon>180, lon-360, lon))

gws_trends_estimates = 
  gws_trends_estimates %>%
  mutate(trends = ifelse(trends < -2, -2, trends)) %>%
  mutate(trends = ifelse(trends > 2, 2, trends)) 

gws.raster = rasterFromXYZ(gws_trends_estimates[,.(lon, lat, trends)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     breaks = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5, 2),
     col = color_pal, 
     legend.args = list(text = 'Trend (cm/yr)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```


## Comparison with In-situ data {.tabset .tabset-pills}

In this section, the downscaled GRACE data was compared with in-situ well data in India.

The main steps involved in the comparison included:

* Linking individual wells to pixels 
* Subsetting the available wells to wells with relatively continuous time-series.
For dug-wells, wells with <20 missing points were included; for bore wells, wells with <50 missing points
were included
* Calculating a mean water-level time-series for each pixels based on the wells in that 
pixel
* Linking the well-based TS with the downscaled GRACE product for the same months
* Estimating the correlation co-efficient between the two

Information about the wells data:

* 28076 wells (bore well + dug wells)
* Available for 4 months of the year (Jan, May, Aug and Nov)
* 1996-2017


### Dug-wells 

```{r echo=FALSE, collapse = TRUE}
#Need to essentially use the gws.raster to create a fishnet polygon
#which will be used to compare the output from GRACE and in-situ
rasterTemplate = gws.raster
vectorTemplate = 
  rasterTemplate %>%
  st_as_stars() %>%
  st_as_sf(as_points = FALSE, crs = "+proj=longlat +datum=WGS84 +no_defs") %>%
  rownames_to_column()

st_crs(vectorTemplate) = "+proj=longlat +datum=WGS84 +no_defs"

#Load the in-situ CGWB data
cgwb = 
  fread('/Users/tejasvi/Dropbox/WB/In-Situ/CGWB_data_wide.csv') %>%
  mutate(Nas = rowSums(is.na(.[,8:91]))) %>%
  dplyr::filter(Nas<20 & SITE_TYPE == 'Dug Well') %>%
  st_as_sf(coords = c("LON", "LAT"), 
           crs = "+proj=longlat +datum=WGS84 +no_defs") %>%
  st_join(vectorTemplate[,c('rowname')]) %>%
  dplyr::select(-V1, -STATE, -DISTRICT, -SITE_TYPE, -Nas)

colname.ref = colnames(cgwb)

cgwb.long = 
  cgwb %>%
  st_drop_geometry() %>% as.data.table() %>%
  melt(id.vars    = c("WLCODE", "rowname"),
    measure.vars  = colname.ref[2:85],
    variable.name = "yearmon",
    value.name    = "measured")

#Now summarize by grid cell
cgwb.grid = 
  cgwb.long %>% 
  group_by(rowname, yearmon) %>%
  dplyr::summarise(measured.mean = mean(measured, na.rm = T)) %>%
  as.data.table() %>%
  mutate(yearmon = as.character(yearmon))

#Link grid-cell id to grace data
gws_05.SA = 
  gws_05.v %>% 
  filter(as.character(REGION_WB) == 'South Asia') %>%
  st_join(vectorTemplate[,c('rowname')]) %>%
  st_drop_geometry() %>% as.data.table() %>%
  data.table::melt(id.vars = c("cell_id", "rowname"),
                   measure.vars = 3:212,
                   variable.name = "ymd", value.name = "GWSA_mean") %>%
  mutate(yearmon = as.yearmon(as.Date(ymd)))

#Create a merged dataset
gws.cgwb.merged = 
  gws_05.SA %>% dplyr::select(rowname, yearmon, ymd, GWSA_mean) %>%
  mutate(yearmon = as.character(yearmon),
         ymd = as.Date(ymd)) %>%
  merge(cgwb.grid, by = c('yearmon', 'rowname')) %>%
  arrange(rowname, ymd)

#Assess the correlation
gws.cgwb.merged.cor = 
  gws.cgwb.merged %>%
  mutate(GWSA_mean = as.numeric(GWSA_mean),
         measured.mean = -as.numeric(measured.mean)) %>%
  group_by(rowname) %>% 
  drop_na() %>%
  dplyr::summarise(cor1=cor(GWSA_mean, measured.mean))
  
#Link the row names to vector
gws.cgwb.cor.v = 
  vectorTemplate %>%
  merge(gws.cgwb.merged.cor, by = 'rowname') %>%
  fasterize(rasterTemplate, field = 'cor1')


color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.cgwb.cor.v,
     breaks = c(-1, -0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1),
     col = color_pal, 
     legend.args = list(text = 'Cor (r)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 
```


### Bore-wells


```{r echo=FALSE, collapse = TRUE}
#Need to essentially use the gws.raster to create a fishnet polygon
#which will be used to compare the output from GRACE and in-situ
rasterTemplate = gws.raster
vectorTemplate = 
  rasterTemplate %>%
  st_as_stars() %>%
  st_as_sf(as_points = FALSE, crs = "+proj=longlat +datum=WGS84 +no_defs") %>%
  rownames_to_column()

st_crs(vectorTemplate) = "+proj=longlat +datum=WGS84 +no_defs"

#Load the in-situ CGWB data
cgwb = 
  fread('/Users/tejasvi/Dropbox/WB/In-Situ/CGWB_data_wide.csv') %>%
  mutate(Nas = rowSums(is.na(.[,8:91]))) %>%
  dplyr::filter(Nas<50 & SITE_TYPE != 'Dug Well') %>%
  st_as_sf(coords = c("LON", "LAT"), 
           crs = "+proj=longlat +datum=WGS84 +no_defs") %>%
  st_join(vectorTemplate[,c('rowname')]) %>%
  dplyr::select(-V1, -STATE, -DISTRICT, -SITE_TYPE, -Nas)

colname.ref = colnames(cgwb)

cgwb.long = 
  cgwb %>%
  st_drop_geometry() %>% as.data.table() %>%
  melt(id.vars    = c("WLCODE", "rowname"),
    measure.vars  = colname.ref[2:85],
    variable.name = "yearmon",
    value.name    = "measured")

#Now summarize by grid cell
cgwb.grid = 
  cgwb.long %>% 
  group_by(rowname, yearmon) %>%
  dplyr::summarise(measured.mean = mean(measured, na.rm = T)) %>%
  as.data.table() %>%
  mutate(yearmon = as.character(yearmon))

#Link grid-cell id to grace data
gws_05.SA = 
  gws_05.v %>% 
  filter(as.character(REGION_WB) == 'South Asia') %>%
  st_join(vectorTemplate[,c('rowname')]) %>%
  st_drop_geometry() %>% as.data.table() %>%
  data.table::melt(id.vars = c("cell_id", "rowname"),
                   measure.vars = 3:212,
                   variable.name = "ymd", value.name = "GWSA_mean") %>%
  mutate(yearmon = as.yearmon(as.Date(ymd)))

#Create a merged dataset
gws.cgwb.merged = 
  gws_05.SA %>% dplyr::select(rowname, yearmon, ymd, GWSA_mean) %>%
  mutate(yearmon = as.character(yearmon),
         ymd = as.Date(ymd)) %>%
  merge(cgwb.grid, by = c('yearmon', 'rowname')) %>%
  arrange(rowname, ymd)

#Assess the correlation
gws.cgwb.merged.cor = 
  gws.cgwb.merged %>%
  mutate(GWSA_mean = as.numeric(GWSA_mean),
         measured.mean = -as.numeric(measured.mean)) %>%
  group_by(rowname) %>% 
  drop_na() %>%
  dplyr::summarise(cor1=cor(GWSA_mean, measured.mean))
  
#Link the row names to vector
gws.cgwb.cor.v = 
  vectorTemplate %>%
  merge(gws.cgwb.merged.cor, by = 'rowname') %>%
  fasterize(rasterTemplate, field = 'cor1')


color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.cgwb.cor.v,
     breaks = c(-1, -0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1),
     col = color_pal, 
     legend.args = list(text = 'Cor (r)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 
```

Results show moderate to poor correlation between the downscaled product and In-Situ wells in:

* For dug wells -> North-Western India
* For bore wells -> peninsular India


## Comparison with non-downscaled GRACE data {.tabset .tabset-pills}

In this section, the downscaled GRACE data was compared with the non-downscaled GRACE products

The main steps involved in the comparison included:

* Disaggregating the 1 degree non-downscaled GRACE grid to 0.5 degrees
* Linking the downscaled and non-downscaled GRACE grids
* Estimating the correlation co-efficient between the downscaled and non-downscaled grids

Information about the non-downscaled GRACE grid:

* 1 degree resolution


```{r echo=FALSE, collapse = TRUE}

#Create raster template
rasterTemplate = 
  rasterFromXYZ(gws_05_wide[,.(lon, lat, ISO_N3)] %>% drop_na())

#Create a vector template
vectorTemplate = 
  rasterTemplate %>%
  st_as_stars() %>%
  st_as_sf(as_points = FALSE, crs = "+proj=longlat +datum=WGS84 +no_defs")

st_crs(vectorTemplate) = "+proj=longlat +datum=WGS84 +no_defs"

vectorTemplate =
  vectorTemplate %>%
  st_join(wb_regions) %>%
  rownames_to_column(var = "rowname_temp") %>%
  rownames_to_column(var = "rowname") %>%
  dplyr::select(-rowname_temp)

#load the non-downscaled GRACE data
gws.OG = 
  fread(paste0(proj_dir, 
               'Outputs/GWS/GRACE_GWS_TWS_Ensemble_SWS_Ensemble_2002_2020_BSL2017.csv')) %>%
  #mutate(lon = lon + 0.001, lat = lat + 0.001) %>% #Jitter so points fall within a single polygon
  st_as_sf(coords = c("lon", "lat"), 
           crs = "+proj=longlat +datum=WGS84 +no_defs") %>%
  st_make_valid() %>%
  st_join(wb_regions) %>%
  #st_join(vectorTemplate[,c('rowname')]) %>%
  drop_na()
  # rownames_to_column(var = "rowname2") #Need a second row ID column as the lower resolution table will have multiples after merging

gws.OG.v =
  vectorTemplate[,c('rowname')] %>%
  st_join(gws.OG, st_nearest_feature)


```

### MENA

```{r echo=FALSE, collapse = TRUE}
template.v = 
  vectorTemplate %>% 
  filter(as.character(REGION_WB) == 'Middle East & North Africa') 

colname.ref = colnames(gws.OG.v)

#Create a long-table with the OG Grace data
gws.OG.sub = 
  gws.OG.v %>%
  filter(as.character(REGION_WB) == 'Middle East & North Africa') %>%
  st_drop_geometry() %>% as.data.table() %>%
  melt(id.vars    = "rowname",
    measure.vars  = colname.ref[3:193],
    variable.name = "yearmon",
    value.name    = "GWSA_OG") %>%
  drop_na()

#Create a long-table with the downscaled GRACE data
gws_05.sub = 
  gws_05.v %>% 
  filter(as.character(REGION_WB) == 'Middle East & North Africa') %>%
  st_join(vectorTemplate[,c('rowname')]) %>%
  st_drop_geometry() %>% as.data.table() %>%
  data.table::melt(id.vars = c("cell_id", "rowname"),
                   measure.vars = 1:212,
                   variable.name = "ymd", value.name = "GWSA_mean") %>%
  mutate(yearmon = substr(as.Date(ymd),1,7)) %>%
  drop_na()


#Create a merged dataset
gws.OG.merged = 
  gws_05.sub %>% dplyr::select(rowname, yearmon, GWSA_mean, ymd, ) %>% 
  merge(gws.OG.sub, by = c('yearmon', 'rowname'), all.x = T) %>%
  arrange(rowname, ymd) %>%
  drop_na()

#Assess the correlation
gws.OG.merged.cor = 
  gws.OG.merged %>%
  mutate(GWSA_mean = as.numeric(GWSA_mean),
         GWSA_OG = as.numeric(GWSA_OG)) %>%
  group_by(rowname) %>% 
  dplyr::summarise(cor1=cor(GWSA_mean, GWSA_OG))
  
#Link the row names to vector
gws.OG.cor.v = 
  template.v %>%
  merge(gws.OG.merged.cor, by = 'rowname') %>%
  drop_na() 

template.r = 
  rasterTemplate %>% 
  crop(gws.OG.cor.v)

gws.OG.cor.v = 
  gws.OG.cor.v %>%
  fasterize(template.r, field = 'cor1')


color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.OG.cor.v,
     breaks = c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1),
     col = color_pal, 
     legend.args = list(text = 'Cor (r)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 
```

### SSA

```{r echo=FALSE, collapse = TRUE}
template.v = 
  vectorTemplate %>% 
  filter(as.character(REGION_WB) == 'Sub-Saharan Africa') 

colname.ref = colnames(gws.OG.v)

#Create a long-table with the OG Grace data
gws.OG.sub = 
  gws.OG.v %>%
  filter(as.character(REGION_WB) == 'Sub-Saharan Africa') %>%
  st_drop_geometry() %>% as.data.table() %>%
  melt(id.vars    = "rowname",
    measure.vars  = colname.ref[3:193],
    variable.name = "yearmon",
    value.name    = "GWSA_OG") %>%
  drop_na()

#Create a long-table with the downscaled GRACE data
gws_05.sub = 
  gws_05.v %>% 
  filter(as.character(REGION_WB) == 'Sub-Saharan Africa') %>%
  st_join(vectorTemplate[,c('rowname')]) %>%
  st_drop_geometry() %>% as.data.table() %>%
  data.table::melt(id.vars = c("cell_id", "rowname"),
                   measure.vars = 1:212,
                   variable.name = "ymd", value.name = "GWSA_mean") %>%
  mutate(yearmon = substr(as.Date(ymd),1,7)) %>%
  drop_na()


#Create a merged dataset
gws.OG.merged = 
  gws_05.sub %>% dplyr::select(rowname, yearmon, GWSA_mean, ymd, ) %>% 
  merge(gws.OG.sub, by = c('yearmon', 'rowname'), all.x = T) %>%
  arrange(rowname, ymd) %>%
  drop_na()

#Assess the correlation
gws.OG.merged.cor = 
  gws.OG.merged %>%
  mutate(GWSA_mean = as.numeric(GWSA_mean),
         GWSA_OG = as.numeric(GWSA_OG)) %>%
  group_by(rowname) %>% 
  dplyr::summarise(cor1=cor(GWSA_mean, GWSA_OG))
  
#Link the row names to vector
gws.OG.cor.v = 
  template.v %>%
  merge(gws.OG.merged.cor, by = 'rowname') %>%
  drop_na() 

template.r = 
  rasterTemplate %>% 
  crop(gws.OG.cor.v)

gws.OG.cor.v = 
  gws.OG.cor.v %>%
  fasterize(template.r, field = 'cor1')


color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.OG.cor.v,
     breaks = c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1),
     col = color_pal, 
     legend.args = list(text = 'Cor (r)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 
```

### South Asia

```{r echo=FALSE, collapse = TRUE}
template.v = 
  vectorTemplate %>% 
  filter(as.character(REGION_WB) == 'South Asia') 

colname.ref = colnames(gws.OG.v)

#Create a long-table with the OG Grace data
gws.OG.sub = 
  gws.OG.v %>%
  filter(as.character(REGION_WB) == 'South Asia') %>%
  st_drop_geometry() %>% as.data.table() %>%
  melt(id.vars    = "rowname",
    measure.vars  = colname.ref[3:193],
    variable.name = "yearmon",
    value.name    = "GWSA_OG") %>%
  drop_na()

#Create a long-table with the downscaled GRACE data
gws_05.sub = 
  gws_05.v %>% 
  filter(as.character(REGION_WB) == 'South Asia') %>%
  st_join(vectorTemplate[,c('rowname')]) %>%
  st_drop_geometry() %>% as.data.table() %>%
  data.table::melt(id.vars = c("cell_id", "rowname"),
                   measure.vars = 1:212,
                   variable.name = "ymd", value.name = "GWSA_mean") %>%
  mutate(yearmon = substr(as.Date(ymd),1,7)) %>%
  drop_na()

#Create a merged dataset
gws.OG.merged = 
  gws_05.sub %>% dplyr::select(rowname, yearmon, GWSA_mean, ymd, ) %>% 
  merge(gws.OG.sub, by = c('yearmon', 'rowname'), all.x = T) %>%
  arrange(rowname, ymd) %>%
  drop_na()

#Assess the correlation
gws.OG.merged.cor = 
  gws.OG.merged %>%
  mutate(GWSA_mean = as.numeric(GWSA_mean),
         GWSA_OG = as.numeric(GWSA_OG)) %>%
  group_by(rowname) %>% 
  dplyr::summarise(cor1=cor(GWSA_mean, GWSA_OG))
  
#Link the row names to vector
gws.OG.cor.v = 
  template.v %>%
  merge(gws.OG.merged.cor, by = 'rowname') %>%
  drop_na() 

template.r = 
  rasterTemplate %>% 
  crop(gws.OG.cor.v)

gws.OG.cor.v = 
  gws.OG.cor.v %>%
  fasterize(template.r, field = 'cor1')


color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.OG.cor.v,
     breaks = c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1),
     col = color_pal, 
     legend.args = list(text = 'Cor (r)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 
```


Results show moderate to poor correlation between the two GRACE products in:

* Yemen/Oman region
* Mouth of the Nile Basin
* Region close to Kuwait
* South Africa/Madagascar/Western Sahara
* Bengal/Indus River Basin/Central India

## Session Info

<div  class="fold s">

```{r echo = FALSE, collapse = TRUE}
print(sessionInfo())
```

</div>