---
title: "Groundwater Depletion Grids"
author: "Tejasvi Hora"
date: "`r Sys.Date()`"
analyst: |
  | Tejasvi Hora
output: 
  html_document:
    df_print: kable
    toc: true
    toc_float: true
    toc_depth: 4
    theme: spacelab
knit: (function(inputFile, encoding) { 
          rmarkdown::render(inputFile,
                            encoding = encoding, 
                            output_file = 'dGWS_binary_assessment_20221030.html',  
                            output_dir = '/Users/tejasvi/Dropbox/WB/GRACE_Ensemble/Scripts/Downscaled/') })
editor_options: 
  chunk_output_type: console
---

## Introduction
This assessment relies on:

* Downscaled GRACE data-provided by __Prof. Di Long, Li Chen and Ximing Wang.__
* Non-downscaled GRACE data processed by Tejasvi Hora

The purpose of this report is to:

* Calculate the GRACE Groundwater Drought Index (GGDI) -- 10.1016/j.rse.2017.06.026
* Create a binary grid highlighting regions with and without depletion


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidync)
library(data.table)
library(tidyverse)
library(lubridate)
library(zyp)
library(sf)
library(raster)
library(terra)
library(RColorBrewer)
library(rasterVis)
library(xts)
library(stars)
library(fasterize)
```

## Load and pre-process the datasets

Main steps involved:

* Loading the dataset: `GWS_mean_05deg.nc`
* Linking points with World Bank regions data

```{r echo = FALSE, collapse = TRUE, cache=TRUE, results='hide'}
proj_dir = "/Users/tejasvi/Dropbox/WB/GRACE_Ensemble/"

pathData = '/Users/tejasvi/Dropbox/gwflagship_GRACEdownscaling/Downscaled TWS_GWS v2/'
gws_mean_05 = 
  tidync(paste0(pathData, 'GWS_mean_05deg.nc')) %>%
  hyper_tibble() %>% as.data.table()

#Create a date sequence
date.seq = 
  seq(as.Date("2003/2/1"), as.Date("2021/9/1"), "month") %>%
  as.data.table() %>%
  rownames_to_column() %>%
  .[, rowname := as.integer(rowname)]
colnames(date.seq) = c('rowname', 'date')

#Merge the dates using the rowname column
gws_mean_05 = 
  gws_mean_05 %>%
  merge(date.seq, by.x = 'time', by.y = 'rowname', all.x = T) %>%
  .[, ':='(yearmon = substr(date, 1, 7),
           year = substr(date, 1, 4),
           month = substr(date, 6, 7),
           GWSA_mean = (GWSA_CLSM+GWSA_Noah)/2,
           grid.id = paste0(lat, lon))] 


####Load World Regions
wb_regions = 
  st_read(paste0(proj_dir, "Spatial Files/WB_Regions/WB_countries_Admin0_10m.shp")) %>%
  dplyr::select(WB_NAME, ISO_A2, ISO_A3, ISO_N3, TYPE, REGION_WB) %>%
  filter(TYPE != 'Dependency') %>%
  st_make_valid()



#load the non-downscaled GRACE data
gws.OG = 
  fread(paste0(proj_dir, 
               'Outputs/GWS/GRACE_GWS_TWS_Ensemble_SWS_Ensemble_2002_2020_BSL2017.csv')) %>%
  #mutate(lon = lon + 0.001, lat = lat + 0.001) %>% #Jitter so points fall within a single polygon
  st_as_sf(coords = c("lon", "lat"), 
           crs = "+proj=longlat +datum=WGS84 +no_defs") %>%
  st_make_valid() %>%
  st_join(wb_regions) %>%
  #st_join(vectorTemplate[,c('rowname')]) %>%
  drop_na()

pathOut =  "/Users/tejasvi/Dropbox/WB/GRACE-Deficit/"

degree.3.r = 
  raster(paste0(pathOut, '3degree_placement.tif'))

degree.3.v = 
  st_as_stars(degree.3.r) %>%
  st_as_sf() %>%
  mutate(grid.id.3 = row_number()) %>%
  dplyr::select(grid.id.3) 


#Link the 3 degree grid id to the GRACE data
gws.OG = 
  gws.OG %>%
  st_join(degree.3.v, join = st_within)

#Pivot longer
#Create 3 degree averages for GWS
gws.OG.df = 
  gws.OG %>%
  st_drop_geometry() %>%
  pivot_longer(cols = `2002-04`:`2021-01`,
               names_to = 'yearmon',
               values_to = 'GWS') %>%
  mutate(year = substr(yearmon, 1, 4),
         month = substr(yearmon, 6, 7))

gws.OG.3 = 
  gws.OG.df %>%
  group_by(year, month, grid.id.3) %>%
  summarise(GWS = mean(GWS, na.rm =  T)) %>%
  as.data.table() 


```


## Trend Analysis (Downscaled) {.tabset .tabset-pills}

*Reference to Previous Analysis*

The steps involved in assessing trends were:

* Conversion of monthly GWS estimates to annual
* Estimating _linear_ trends by estimating Sen's Slope
* Significance of trend not assessed

```{r echo=FALSE, collapse = TRUE}
#Convert to wide
gws_05_wide = 
  gws_mean_05 %>% 
  dplyr::select(GWSA_mean, lat, lon, date) %>%
  tidyr::spread(key = date, value = GWSA_mean) %>%
  mutate(cell_id = paste0(lon, lat))

#Convert to spatial object
gws_05.v = 
  st_as_sf(gws_05_wide, coords = c("lon", "lat"), 
           crs = "+proj=longlat +datum=WGS84 +no_defs")


#Merge country data with GRACE
gws_05.v = 
  gws_05.v %>%
  st_make_valid() %>%
  st_join(wb_regions)

#Add the WB regions to the data.table version
gws_05_wide = 
  gws_05_wide %>%
  merge(gws_05.v[,c(213:219)] %>% st_drop_geometry(), by = 'cell_id', all.x = T)


```


### MENA 

```{r echo=FALSE, collapse = TRUE}
#Conduct a trend analysis on each pixel

#First, convert values from monthly to year
gws_annual = 
  gws_05_wide %>% 
  .[as.character(REGION_WB) == 'Middle East & North Africa'] %>%
  data.table::melt(id.vars = c("lon", "lat", "cell_id"),
                   measure.vars = 3:214,
                   variable.name = "ymd", value.name = "GWSA_mean") %>% 
  .[, year := substr(ymd, 1,4)] %>%
  .[, an_mean := mean(GWSA_mean), .(lat, lon, year)] %>% 
  .[,.(lat, lon, cell_id, year, an_mean)] %>% distinct() %>%
  dcast(lon + lat + cell_id ~ year, value.var = "an_mean") %>% .[order(cell_id)]

gws_trends = 
  apply(gws_annual[, 4:ncol(gws_annual)], 1, zyp.yuepilon)

gws_trends_estimates = 
  gws_annual[,1:3] %>%
  mutate(trends = gws_trends[2,], sig = gws_trends[6,]) %>%
  mutate(lon = ifelse(lon>180, lon-360, lon))

gws_trends_estimates = 
  gws_trends_estimates %>%
  mutate(trends = ifelse(trends < -2, -2, trends)) %>%
  mutate(trends = ifelse(trends > 2, 2, trends)) 

gws.raster = rasterFromXYZ(gws_trends_estimates[,.(lon, lat, trends)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     breaks = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5, 2),
     col = color_pal, 
     legend.args = list(text = 'Trend (cm/yr)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```

### SSA

```{r echo=FALSE, collapse = TRUE}

#First, convert values from monthly to year
gws_annual = 
  gws_05_wide %>% 
  .[as.character(REGION_WB) == 'Sub-Saharan Africa'] %>%
  data.table::melt(id.vars = c("lon", "lat", "cell_id"),
                   measure.vars = 3:214,
                   variable.name = "ymd", value.name = "GWSA_mean") %>% 
  .[, year := substr(ymd, 1,4)] %>%
  .[, an_mean := mean(GWSA_mean), .(lat, lon, year)] %>% 
  .[,.(lat, lon, cell_id, year, an_mean)] %>% distinct() %>%
  dcast(lon + lat + cell_id ~ year, value.var = "an_mean") %>% .[order(cell_id)]

gws_trends = 
  apply(gws_annual[, 4:ncol(gws_annual)], 1, zyp.yuepilon)

gws_trends_estimates = 
  gws_annual[,1:3] %>%
  mutate(trends = gws_trends[2,], sig = gws_trends[6,]) %>%
  mutate(lon = ifelse(lon>180, lon-360, lon))

gws_trends_estimates = 
  gws_trends_estimates %>%
  mutate(trends = ifelse(trends < -2, -2, trends)) %>%
  mutate(trends = ifelse(trends > 2, 2, trends)) 

gws.raster = rasterFromXYZ(gws_trends_estimates[,.(lon, lat, trends)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     breaks = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5, 2),
     col = color_pal, 
     legend.args = list(text = 'Trend (cm/yr)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```



### SA

```{r echo=FALSE}

#First, convert values from monthly to year
gws_annual = 
  gws_05_wide %>% 
  .[as.character(REGION_WB) == 'South Asia'] %>%
  data.table::melt(id.vars = c("lon", "lat", "cell_id"),
                   measure.vars = 3:214,
                   variable.name = "ymd", value.name = "GWSA_mean") %>% 
  .[, year := substr(ymd, 1,4)] %>%
  .[, an_mean := mean(GWSA_mean), .(lat, lon, year)] %>% 
  .[,.(lat, lon, cell_id, year, an_mean)] %>% distinct() %>%
  dcast(lon + lat + cell_id ~ year, value.var = "an_mean") %>% .[order(cell_id)]

gws_trends = 
  apply(gws_annual[, 4:ncol(gws_annual)], 1, zyp.yuepilon)

gws_trends_estimates = 
  gws_annual[,1:3] %>%
  mutate(trends = gws_trends[2,], sig = gws_trends[6,]) %>%
  mutate(lon = ifelse(lon>180, lon-360, lon))

gws_trends_estimates = 
  gws_trends_estimates %>%
  mutate(trends = ifelse(trends < -2, -2, trends)) %>%
  mutate(trends = ifelse(trends > 2, 2, trends)) 

gws.raster = rasterFromXYZ(gws_trends_estimates[,.(lon, lat, trends)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     breaks = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5, 2),
     col = color_pal, 
     legend.args = list(text = 'Trend (cm/yr)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```



## Groundwater Deficit Indicator (Downscaled) {.tabset .tabset-pills}

The main steps involved in estimating this indicator were:

* Calculate the month-specific means over the entire time period (2003-2020) (i.e. the climatology)
* Remove the seasonality by substracting the climatology from the month-specifc values
* Normalize the remaining signal by subtracting the mean and dividing by SD (Groundwater deficit)
* Filter to include only 2019 values, and take the mean Groundwater deficit (last 2 years of the dataset)
* Convert to binary based on if a grid value is greater or less than -1.5 (standard deviations from the mean)

**See the paper here for more details: https://www.sciencedirect.com/science/article/abs/pii/S0034425717302869**

Reading the map:

* Red are regions with values = 1 -- regions where GW storage has declined based on the indicator
* Blue are regions with values = 0 -- regions where GW storage has not declined based on the indicator


```{r echo=FALSE, collapse = TRUE}
  
#GGDI steps
GGDI = 
  gws_mean_05 %>%
  filter(year<2021) %>%
  group_by(grid.id, month) %>%
  mutate(GWS.month.mean = mean(GWSA_mean, na.rm = T)) %>% 
  mutate(GWS.climatology = GWSA_mean - GWS.month.mean) %>%
  group_by(grid.id) %>%
  mutate(grid.mean = mean(GWS.climatology, na.rm = T),
         grid.sd = sd(GWS.climatology, na.rm = T)) %>%
  mutate(GWS.deficit = (GWS.climatology-grid.mean)/grid.sd) %>%
  as.data.table()


GGDI.binary = 
  GGDI %>%
  filter(year>=2019) %>%
  dplyr::select(lat, lon, grid.id, GWS.deficit) %>%
  group_by(lat, lon, grid.id) %>%
  summarise(GWS.deficit.mean = mean(GWS.deficit, na.rm = T)) %>% 
  mutate(GWS.binary = ifelse(GWS.deficit.mean > -1.5, 0, 1)) %>%
  as.data.table()


#Convert to spatial object
GGDI.binary.v = 
  st_as_sf(GGDI.binary, coords = c("lon", "lat"), 
           crs = "+proj=longlat +datum=WGS84 +no_defs")

#Merge country data with GRACE
GGDI.binary.v = 
  GGDI.binary.v %>%
  st_make_valid() %>%
  st_join(wb_regions)


GGDI.out = 
  GGDI.binary.v %>% st_drop_geometry() %>% as.data.table() %>% 
  merge(GGDI.binary %>% dplyr::select(lat, lon, grid.id), by = 'grid.id')

pathOut =  "/Users/tejasvi/Dropbox/WB/GRACE-Deficit/"

fwrite(GGDI.out, paste0(pathOut, 'GWS_Deficit_downscaled.csv'))

```

### Eg. No depletion

**Check values towards the end of the time series** 

```{r echo=FALSE, collapse = TRUE}


plot.ts = 
  GGDI[grid.id=='-34.7519.75'] %>%
  mutate(yearmon = as.yearmon(yearmon)) 

plot.zoo = zoo(plot.ts$GWS.deficit, plot.ts$yearmon)

plot(plot.zoo, type='l', ylab = 'Deficit Indicator')

```


### Eg. With depletion

**Check values towards the end of the time series**

```{r echo=FALSE, collapse = TRUE}

plot.ts = 
  GGDI[grid.id=='-30.7526.75'] %>%
  mutate(yearmon = as.yearmon(yearmon)) 

plot.zoo = zoo(plot.ts$GWS.deficit, plot.ts$yearmon)

plot(plot.zoo, type='l', ylab = 'Deficit Indicator')

```


### MENA 

```{r echo=FALSE, collapse = TRUE}

GGDI.region = 
  GGDI.binary.v %>% st_drop_geometry() %>% as.data.table() %>% 
  .[as.character(REGION_WB) == 'Middle East & North Africa'] %>%
  merge(GGDI.binary %>% dplyr::select(lat, lon, grid.id), by = 'grid.id')

gws.raster = rasterFromXYZ(GGDI.region[,.(lon, lat, GWS.binary)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     col = c('#4393c3', '#d6604d'), 
     legend.args = list(text = 'Depletion (yes/no)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```

### SSA

```{r echo=FALSE, collapse = TRUE}

GGDI.region = 
  GGDI.binary.v %>% st_drop_geometry() %>% as.data.table() %>% 
  .[as.character(REGION_WB) == 'Sub-Saharan Africa'] %>%
  merge(GGDI.binary %>% dplyr::select(lat, lon, grid.id), by = 'grid.id')

gws.raster = rasterFromXYZ(GGDI.region[,.(lon, lat, GWS.binary)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     col = c('#4393c3', '#d6604d'), 
     legend.args = list(text = 'Depletion (yes/no)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```



### SA

```{r echo=FALSE}

GGDI.region = 
  GGDI.binary.v %>% st_drop_geometry() %>% as.data.table() %>% 
  .[as.character(REGION_WB) == 'South Asia'] %>%
  merge(GGDI.binary %>% dplyr::select(lat, lon, grid.id), by = 'grid.id')

gws.raster = rasterFromXYZ(GGDI.region[,.(lon, lat, GWS.binary)])

color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(gws.raster,
     col = c('#4393c3', '#d6604d'), 
     legend.args = list(text = 'Depletion (yes/no)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```

## Groundwater Deficit Indicator (Non-Downscaled) {.tabset .tabset-pills}

* Follows the same methodology as the above.
* Uses the 3 by 3 degrees GRACE grid 

Reading the map:

* Red are regions with values = 1 -- regions where GW storage has declined based on the indicator
* Blue are regions with values = 0 -- regions where GW storage has not declined based on the indicator


```{r echo=FALSE, collapse = TRUE}
 
#GGDI steps
GGDI = 
  gws.OG.3 %>%
  filter(year<2021) %>%
  group_by(grid.id.3, month) %>%
  mutate(GWS.month.mean = mean(GWS, na.rm = T)) %>% 
  mutate(GWS.climatology = GWS - GWS.month.mean) %>%
  group_by(grid.id.3) %>%
  mutate(grid.mean = mean(GWS.climatology, na.rm = T),
         grid.sd = sd(GWS.climatology, na.rm = T)) %>%
  mutate(GWS.deficit = (GWS.climatology-grid.mean)/grid.sd) %>%
  as.data.table()


GGDI.binary = 
  GGDI %>%
  filter(year>=2019) %>%
  dplyr::select( grid.id.3, GWS.deficit) %>%
  group_by(grid.id.3) %>%
  summarise(GWS.deficit.mean = mean(GWS.deficit, na.rm = T)) %>% 
  mutate(GWS.binary = ifelse(GWS.deficit.mean > -1.5, 0, 1)) %>%
  as.data.table()


#Convert to spatial object
GGDI.binary.v = 
  degree.3.v %>%
  merge(GGDI.binary, by = 'grid.id.3') 

GGDI.binary.centroid = 
  GGDI.binary.v %>% 
  st_centroid() %>%
  st_transform(st_crs(wb_regions)) %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry() %>%
  as.data.table()

pathOut =  "/Users/tejasvi/Dropbox/WB/GRACE-Deficit/"

fwrite(GGDI.binary.centroid, paste0(pathOut, 'GWS_Deficit_3degree.csv'))

st_write(GGDI.binary.v, 
         paste0(pathOut, 'GWS_Deficit_3degree.shp'),
         delete_layer = T)

# 
# 
# #Merge country data with GRACE
# GGDI.binary.v = 
#   GGDI.binary.v %>%
#   st_make_valid() %>%
#   st_join(wb_regions)
# 
# 
# GGDI.out = 
#   GGDI.binary.v %>% st_drop_geometry() %>% as.data.table() %>% 
#   merge(GGDI.binary %>% dplyr::select(lat, lon, grid.id), by = 'grid.id')
# 
# pathOut =  "/Users/tejasvi/Dropbox/WB/GRACE-Deficit/"
# 
# fwrite(GGDI.out, paste0(pathOut, 'GWS_Deficit_downscaled.csv'))


```

### Eg. With depletion

**Check values towards the end of the time series**

```{r echo=FALSE, collapse = TRUE}

plot.ts = 
  GGDI[grid.id.3==1158] %>%
  mutate(yearmon = as.yearmon(paste0(year,'-', month))) 

plot.zoo = zoo(plot.ts$GWS.deficit, plot.ts$yearmon)

plot(plot.zoo, type='l', ylab = 'Deficit Indicator')

```


### Global Hotspots 

```{r echo=FALSE, collapse = TRUE}


GGDI.binary.r = 
  fasterize(GGDI.binary.v, degree.3.r, field = 'GWS.binary')

GGDI.binary.r = 
  crop(GGDI.binary.r, extent(-180, 180, -60, 60))


color_pal = c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

plot(GGDI.binary.r,
     col = c('#4393c3', '#d6604d'), 
     legend.args = list(text = 'Depletion (yes/no)', side = 4, 
                        font = 2, line = 2.5, cex = 0.8)) 

```


## Session Info

<div  class="fold s">

```{r echo = FALSE, collapse = TRUE}
print(sessionInfo())
```

</div>